Data Deduplication and Zipf-like Distribution in VM Cloud Storage

### 2012-03-01 vm_snapshot_sim/2012arch.tex
Tao

At the runtime, the CDS index resides in a distributed  memory lookup table  
implemented using Memcached~\cite{memcached} to leverage the aggregated memory in a cluster.
Th usage of memory for each machine is small and thus  this scheme  does not
lead to  a memory resource contention with the existing cloud services.
For example, in our current implementation setting at Aliyun, we use up to  150MB memory per node during night backup
including CDS and other metadata for all levels of deduplication. Most of memory resource (48GB per node 
in our experiment) is used for existing  applications.
Each CDS content data stored  in the distributed file system
has multiple copies in different machines for the purpose of fault tolerance and 
while providing high read  throughput.  
%The data of CDS is stored in a special block store in the distribued file system.

@@

Thus our CDS selection strategy is designed as follows.
\begin{itemize}
\item First, most popular items from OS disks are selected.
Our experiment data shows that about 
100GB data ($25\times 10^6$\cite{BAD}) from 7 most popular OS releases
can cover over 70\% of OS snapshots and the index of these 100GB OS data blocks
only takes about total 800MB in the CDS index and then memory requirement per node for running the backup
on  100 nodes is tiny.
We expect in practice the OS CDS index may grow up to 2GB in order to cover large varieties  of commonly deployed
OS releases in practice.
\item Second, the degree of duplication is computed for snapshot blocks of data disks and
% and also snapshot 
%blocks of user data items stored in the OS disk (some users store data in their OS disks).
the  most popular items are selected. As discussed in Section~\ref{sect:exper}, 
about $10^7$ (???) most popular items
can cover about 37\% of content blocks appeared in various snapshots after level 1 and level deduplication.  
\end{itemize}

>>>>

At the runtime, the CDS index resides in a distributed  memory lookup table  
implemented using Memcached~\cite{memcached} to leverage the aggregated memory in a cluster.
Th usage of memory at each machine is small and thus  this scheme  does not
lead to  a memory resource contention with the existing cloud services.
For example, in our current implementation setting at Aliyun, we use up to  150MB memory per node during night backup
including CDS and other metadata for all levels of deduplication. Most of memory resource (48GB per node 
in our experiment) is used for existing  applications.
Each CDS content data stored  in the distributed file system
has multiple copies in different machines for the purpose of fault tolerance and 
while providing high read  throughput.  
%The data of CDS is stored in a special block store in the distribued file system.

@@

As the number of VMs increases, the popularity of common blocks for data disks changes.
But this increase has a small impact on OS block popularity because 
there are a limited number of OS releases used by VMs. We let CDS cover the common OS blocks first and
our CDS selection strategy is designed as follows.
\begin{itemize}
\item First, most popular items from OS disks are selected.
Our experiment data shows that about 
100GB data from 7 most popular OS releases
can at least cover over 70\% of OS snapshots and the index of these 100GB OS blocks
only takes about total 1GB in the CDS index and then memory requirement per node for running the backup
on  a large cluster is tiny.
We expect that the OS CDS index may grow up to 2GB in order to cover large varieties  of commonly deployed
OS releases in practice.
\item Second, the degree of duplication is computed for snapshot blocks of data disks and
% and also snapshot 
%blocks of user data items stored in the OS disk (some users store data in their OS disks).
the  most popular items are selected. As discussed in Section~\ref{sect:exper}, 
that about 1\% of data blocks are selected based on their duplicate counts and these popular blocks
can cover about 38\% of data  blocks appeared in various snapshots after level 1 and level deduplication.  
The memory size per node used for CDS is within 150MB.
\end{itemize}
Thus in total the CDS index size per node takes less than 160MB in a large cluster bigger than 100 nodes based on
Aliyun's experiment data, covering over 53\% of blocks in OS and data disks after inner VM deduplication.
This memory usage is well below the limit required by Aliyun.

### 2012-03-01 vm_snapshot_sim/2012background.tex
Tao

On the other hand, the computing resource allocated for backup is often very small.
This is because snapshot data is not retrieved frequently by users and the allocation of  cluster resource
is typically prioritized towards active services such as online web services and map-reduce jobs.
Thus  it is not easy to accept any backup behavior that affects VMs performance or stability of active users.

%snapshot service shall not compete cpu, memory, or I/O bandwidth with VMs. Specifically, the memory usage of snapshot service can never exceed 500MB in any node.
\item {\em Fast backup speed.}
The snapshot  backup  operations need to be completed quickly  every day (e.g. 1-2 hours ). 
Often a cloud service has  few hours of light workload each day (e.g. midnight),  which creates an opportunity  for backup.
But a  longer use of bandwidth and computing resource  for backup can create  noticeable  contention with the existing cloud,
which is not preferred for cloud production system operation. 
%As a large-scale cluster hosting tens of thousand of active VMs everyday,  the amount of data
%to be processed is huge. 
%For example, in an Aliyun cluster with over 1,000 nodes and each hosts over 25 VMs, The aggreated amount of data 
 % the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.

@@

There are multiple choices in designing a backup architecture  for VM snapshots.
We discuss the following design options with a consideration on their strengths and weakness.
\begin{enumerate}
\item  {\em An external and dedicated backup storage system.} 
In this architecture setting, a separate backup storage system using
the standard backup and deduplication techniques can be deployed~\cite{bottleneck08,extreme_binning09,sparseindex09}. 
This system is attached to the cloud network and every machine can periodically transfer snapshot data to 
the attached backup system. 
A key weakness of this approach is communication bottleneck between a large number of machines
in a cloud to this centralized  service.
Another weakness is that the cost of allocating separate resource for dedicated backup  can be expensive.
Since most of backup data is not used eventually, CPU and memory resource in such a backup cluster may not be fully utilized.
\item {\em A decentralized and co-hosted backup system with full deduplication.}
In this option, the backup system runs on an existing set of cluster machines. 
The advantage is that there is no separate and dedicated cost for backup.
% and a distributed storage architecture for backup allows a possible exploitation of  data locality between
%the source of data and storage location of backup data. 
The disadvantage is that 
%the backup service would compete CPU, memory, and disk resources with the other cloud services.
even such a  backup service may only use  a fraction of the existing disk storage, 
fingerprint-based search does require a significant amount of memory for fast lookup of identical content chunks.
This competes memory resource  with the existing applications.
%Decentralized deduplication is studied in ~\cite{Clements2009} and the focus is on
%block-level copy-on-write and compare-by-value techniques.
Even approximation~\cite{extreme_binning09,sparseindex09} can be used to reduce memory requirement,  
a  key weakness of global content fingerprint comparison is that it can affect fault isolation.
Because a content chunk is compared with a content signature collected from other users,
this artificially creates data dependence among different cloud users.
% since storage for shared identical content chunks can become a failure point.
The loss of a common chunk can affect many users whose VM images share this 
chunk. Adding replicas for this global dataset can enhance the availability,
but incurs significantly more cost.
%2) data sharing among users for accessing common signagutes causes the system less resilient to storage failures.
%any loss of one piece of  shared data content hash and actualcontent will damange many VM snapshots, which can cause massive impacts
%on reliability of many users.

>>>>

On the other hand, the computing resource allocated for backup is often very small.
This is because snapshot data is not retrieved frequently by users and the allocation of  cluster resource
is typically prioritized towards active services such as online web services and map-reduce jobs.
Thus  it is not easy to accept any backup behavior that affects VMs performance or stability of active users.  
At Aliyun, it is required that while CPU and disk usage should be  small or modest during backup time,
the memory footprint of deduplication should not exceed 1\% of memory at each node, which is about 480MB.

%snapshot service shall not compete cpu, memory, or I/O bandwidth with VMs. Specifically, the memory usage of snapshot service can never exceed 500MB in any node.
\item {\em Fast backup speed.}
The snapshot  backup  operations need to be completed quickly  every day (e.g. 1-2 hours ). 
Often a cloud service has  few hours of light workload each day (e.g. midnight),  which creates an opportunity  for backup.
But a  longer use of bandwidth and computing resource  for backup can create  noticeable  contention with the existing cloud,
which is not preferred for cloud production system operation. Thus it is desirable that backup for all nodes
can be conducted in parallel and any centralized or  cross-machine communication for
deduplication should not become a bottleneck.
%As a large-scale cluster hosting tens of thousand of active VMs everyday,  the amount of data
%to be processed is huge. 
%For example, in an Aliyun cluster with over 1,000 nodes and each hosts over 25 VMs, The aggreated amount of data 
 % the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.

@@

There are multiple choices in designing a backup architecture  for VM snapshots.
We discuss the following design options with a consideration on their strengths and weakness.
\begin{enumerate}
\item  {\em An external and dedicated backup storage system.} 
In this architecture setting, a separate backup storage system using
the standard backup and deduplication techniques can be deployed~\cite{bottleneck08,extreme_binning09,sparseindex09}. 
This system is attached to the cloud network and every machine can periodically transfer snapshot data to 
the attached backup system. 
A key weakness of this approach is communication bottleneck between a large number of machines
in a cloud to this centralized  service.
Another weakness is that the cost of allocating separate resource for dedicated backup  can be expensive.
Since most of backup data is not used eventually, CPU and memory resource in such a backup cluster may not be fully utilized.
\item {\em A decentralized and co-hosted backup system with full deduplication.}
In this option, the backup system runs on an existing set of cluster machines. 
The advantage is that there is no separate and dedicated cost for backup.
% and a distributed storage architecture for backup allows a possible exploitation of  data locality between
%the source of data and storage location of backup data. 
The disadvantage is that 
%the backup service would compete CPU, memory, and disk resources with the other cloud services.
even such a  backup service may only use  a fraction of the existing disk storage, 
fingerprint-based search does require a significant amount of memory for fast lookup of identical content chunks.
This competes memory resource  with the existing applications.
%Decentralized deduplication is studied in ~\cite{Clements2009} and the focus is on
%block-level copy-on-write and compare-by-value techniques.

Even approximation~\cite{extreme_binning09,sparseindex09} can be used to reduce memory requirement,  
a  key weakness of global content fingerprint comparison is that it can affect fault isolation.
Because a content chunk is compared with a content signature collected from other users,
this artificially creates data dependence among different cloud users.
% since storage for shared identical content chunks can become a failure point.
The loss of a common chunk can affect many users whose VM images share this 
chunk. Adding replicas for this global dataset can enhance the availability,
but incurs significantly more cost.
%2) data sharing among users for accessing common signagutes causes the system less resilient to storage failures.
%any loss of one piece of  shared data content hash and actualcontent will damange many VM snapshots, which can cause massive impacts
%on reliability of many users.

### 2012-03-01 vm_snapshot_sim/2012eval.tex
Tao

Figure~\ref{fig:overall} shows the overall impact of 3-level deduplication.
The X axis shows the overall impact in (a),  impact on OS disk in (b), and impact on data disk in (c).
Each bar in the Y axis shows the data size after deduplication divided by the original data size.
Level-1 elimination can reduce the data size to 22\% of original data, namely it delivers close 80\% reduction.
Level-2 elimination applied after level 1
reduces the size further to 16\% of original size, namely it delivers additional 6\% reduction.
Level-3 elimination together with level 1 and 2
reduces the size further to 9\% of original size, namely it delivers additional 7\% reduction.
Level 2 elimination is more visible in OS disk than data disk, because data change frequency is really small
when we sample last 10 snapshots of each user in 10 days. Nevertheless, the average impact of level 2 is also significant.
A 6-7\% of reduction from the original data represents about 600-700TB of space.

@@

Overall speaking, inner   VM deduplication or  CDS-based deduplication
can work well alone, but by combining them together we get a fairly good and stable deduplication ratio to 
all kind of OSes. 
Compared to a traditional dirty bit approach based on pages of
each file (e.g. segment in our scheme),
our CDS-based level 3 approach  can save additional 50\% storage space because many of level 2 block
content can be eliminated using the CDS also.

@@

Figure~\ref{fig:OSdatacoverage} shows the cumulative coverage  ratio of  popular common OS blocks. 
From both graphs, a small number of most popular common blocks can coverage a large percentage of 
snapshot blocks and thus the size of CDS can be chosen to be small while still having a large
coverage. 
% with $\alpha$ close to 0.78.
%A tiny hottest fraction of data lead to majority of data duplication, as a result, we can target at this small set of commonly seen data
%to design our fast and lightweight deduplication process, just like what people did in CDN and web proxy.

\begin{figure}
\centering
%\epsfig{file=images/cdf.os.v4.eps, width=3in}
%\epsfig{file=images/cdf.os.v4.eps, width=3in}
\epsfig{file=images/ay41a_small.os_incremental.cdf.eps, width=3in}
\caption{ Cumulative  coverage of popular common OS disk blocks.}
\label{fig:OSdatacoverage}
\end{figure}

@@

The selection of CDS is completed by separating OS disk and data disk.
While block commonality among data disks may be affected in some degree
 when we scale up the number of total VMs  to be supported,
we expect that  the block commonality among OS disks still remains to be high and  
can be well covered by a fixed CDS because users still adopt common OS releases.
To  verify this,  we examine the percentage of common  OS blocks of the same
release in different VMs  OS releases in Figure~\ref{fig:OSunchanged}. 
%we choose 7 major OSes in Aliyun's
%VM cloud platform and they are: 
%Win2008 Server 64 bits, Win2003 Server 32 bits, Win2003 Server 64 bits, RHEL, CentOS, Ubuntu Server and Debian (all Linux
%distributions are 64 bits).
%we examine 5 VM user disks from each OS, and 10 snapshots for each VM. These VMs have been actively used by their
%owners. Some of  OS disks are modified frequently and in some cases,  users even store a large amount of user data on
%their OS disks. 
%For example, some Ubuntu OS disks have only 2GB of data while some have 50GB.
%For each OS, the base image and all OS
%disk snapshots are divided into variable-sized blocks. Then 
We consider the earlier snapshot for each OS disk as the base image.
For every block in the base image, 
we classify this block as ``completely common''
if  this block has appeared in all snapshots of the same OS release even they are used by different VMs.
Figure \ref{fig:OSunchanged} shows that for each
OS,  at least 70\% of OS blocks are common among all snapshots of the same OS. 
Some latest release  of OS tends to have a higher percentage of content change
while  old release tends to have more variations of content versions.
That can be interpreted as that users with very old  version of operating systems
have a lower tendency to update their OS versions and this causes a larger discrepancy
among OS snapshots of these users.

@@

????????????????????????????????????????????????
Based on the above analysis, we have selected a small set of most popular
OS blocks, which is 1.5\% of  unique OS blocks. This set
is quite sufficient to cover the OS related blocks. 
%Then we use this CDS to run the deduplication process again.
That correspond to about  80GB of OS blocks and its corresponding CDS meta occupies 800MB in the cache.
%  data from 350 OS disk snapshots,

>>>>

Figure~\ref{fig:overall} shows the overall impact of 3-level deduplication.
The X axis shows the overall impact in (a),  impact on OS disk in (b), and impact on data disk in (c).
Each bar in the Y axis shows the data size after deduplication divided by the original data size.
Level-1 elimination can reduce the data size to about 22\% of original data, namely it delivers close 78\% reduction.
Level-2 elimination applied after level 1
reduces the size further to about 18.5\% of original size, namely it delivers additional 4.5\% reduction.
Level-3 elimination together with level 1 and 2
reduces the size further to 8\% of original size, namely it delivers additional 10.5\% reduction.
Level 2 elimination is more visible in OS disk than data disk, because data change frequency is really small
when we sample last 10 snapshots of each user in 10 days. Nevertheless, the overall impact of level 2 is still significant.
A 4.5\% of reduction from the original data represents about 450TB space saving for a 1000-node cluster.

@@

%Overall speaking, inner   VM deduplication or  CDS-based deduplication
%can work well alone, but by combining them together we get a fairly good and stable deduplication ratio to 
%all kind of OSes. 
%Compared to a traditional dirty bit approach based on pages of
%each file (e.g. segment in our scheme),
%our CDS-based level 3 approach  can save additional 50\% storage space because many of level 2 block
%content can be eliminated using the CDS also.

@@

%Figure~\ref{fig:OSdatacoverage} shows the cumulative coverage  ratio of  popular common OS blocks. 
%From both graphs, a small number of most popular common blocks can coverage a large percentage of 
%snapshot blocks and thus the size of CDS can be chosen to be small while still having a large
%coverage.

%\begin{figure}
%\centering
%\epsfig{file=images/ay41a_small.os_incremental.cdf.eps, width=3in}
%\caption{ Cumulative  coverage of popular common OS disk blocks.}
%\label{fig:OSdatacoverage}
%\end{figure}

@@

Because there are a limited number of OS releases supported by Aliyun, we study common blocks  among
OS disks loaded with  the same OS release. 
%The selection of CDS is completed by separating OS disk and data disk.
%While block commonality among data disks may be affected in some degree
% when we scale up the number of total VMs  to be supported,
%we expect that  the block commonality among OS disks still remains to be high and  
%%can be well covered by a fixed CDS because users still adopt common OS releases.
%we choose 7 major OSes in Aliyun's
%VM cloud platform and they are: 
%Win2008 Server 64 bits, Win2003 Server 32 bits, Win2003 Server 64 bits, RHEL, CentOS, Ubuntu Server and Debian (all Linux
%distributions are 64 bits).
%we examine 5 VM user disks from each OS, and 10 snapshots for each VM. These VMs have been actively used by their
%owners. Some of  OS disks are modified frequently and in some cases,  users even store a large amount of user data on
%their OS disks. 
%For example, some Ubuntu OS disks have only 2GB of data while some have 50GB.
%For each OS, the base image and all OS
%disk snapshots are divided into variable-sized blocks. Then 
In  Figure~\ref{fig:OSunchanged},  we list a conservative  commonality study in 7 major OS versions  supported
in Aliyun.  We consider the earlier snapshot for each OS disk as the base image.
For every block in the base image, we classify this block as ``completely common''
if  this block has appeared in all snapshots of the same OS release even they are used by different VMs.
Figure \ref{fig:OSunchanged} shows that for each
OS,  at least 70\% of OS blocks are completely common among all snapshots of the same OS. 
Some latest release  of OS tends to have a higher percentage of content change
while  old release tends to have more variations of content versions.
That can be interpreted as that users with very old  version of operating systems
have a lower tendency to update their OS versions and this causes a larger discrepancy
among OS snapshots of these users.

@@

Based on the above analysis, we have selected a small set of most popular
OS blocks, which is about 100GB OS data and its corresponding CDS index takes about 1GB memory space in total.
They can cover sufficiently over 70\% OS blocks.

%Then we use this CDS to run the deduplication process again.
%That correspond to about  80GB of OS blocks and its corresponding CDS meta occupies 800MB in the cache.
%  data from 350 OS disk snapshots,
